Exercise 3: Software Architecture in Practice Case: Beer Web Store

### 2008-03-17 trunk/SAiP/Module1/H3/doc/qualityattributes.tex
Updating with spell-check

Consider the following main use case for a Beer Web Store for specialty
beers:

@@

Naturally in order for the scenarios to be able to be used to clarify the architecture it must be possible quantify them, and place them in different categories. This requires that the above scenario(s) are extended with e.g. response time, how many times the user should be able to perform an operation before experiencing a fault (down-town).

Once the scenarios are clarified and voted on, then the top priority scenarios may be formalized and it may be determined which quality attribute(s) they involve. For examples sake lets asume that the above scenario was selected as top priority.

\fixme{Insert table with quality attribute breakdown}

As it may be seen, the scenario(s) focus primarily on \emph{Availability} (how rare should it be that a user experience non-resonsiveness and how many simultanious users can the system support), \emph{Usability} (how easy should it be for the user to cancel a transaction halfway through and how easy should it be to pick up a previous transaction), \emph{Performance} (how big a delay can the user expect when performing an operation like a lookup in the database), \emph{Security} (how to avoid someone using stolen credit cards and also prevent a "hacker" from stealing the credit card information either durig the transaction or from the Beer Web Store's database.

@@

The scenarios may be further clarified in order to include the aspects of availability and performance, as shown below.

\begin{quote}
The customer browses the beer catalogue for descriptions and reviews and receives the result within 5 seconds 99\% of the time, 30 seconds 0,9\% of the time and more than 30 seconds 0,1\% of the time.

The customer selects appropriate beers for purchase and these are added to his shopping basket. When the customer checks out, payment is validated through the payment service within 30 seconds.
\end{quote}

\fixme{Answer question 2.2}

\subsection{Quality Attributes - Question 3}

@@

Finding out if a resource is down may be done using one of the fault detection tactice; ping/echo, heartbeat or exception. Due to the loose coupling between the client (browser) and the server, the exception solution is not benificial. Since it may be assumed that the supervising process (remotely deployed) (the process monitoring if the server is down) is stable, then a heartbeat is sufficient to ensure availability, meaning that the web-server will issue a heart-beat every e.g. 1 second to a maintance process on the same or another machine, and if one (or perhaps more) heartbeats is missed then the maintanance service can act.

The proper response to a faulty server may be many things, as defined by the tactics group \emph{fault recovery}. As the requirement for availability stipulated a three stage fault recovery, the server or servers (as it may be required to have multiple serves to do load-balancing, depending on how many users must be supported) can service the 5 second requirement, and tehn if an error occur (as detected by the heartbeat) then restart or swap to another server/database in 30 seconds. As 30 seconds is insufficient to bott a webserver and a database, the Spare tactic would be insufficient. At the same tiem having Active redundancy would be over-kill, as a this kind of immidiat switch-over is not required. For these reasons Passive Redundancy is sufficient.

As it is an internet system, it will also be sufficient to simply restart the session, and no \emph{Shadow}, \emph{State resynchronization} or \emph{Rollback} is needed in the main web server. The user may experience beeing thrown back to the main page, yet this is deemed acceptable, under availability and performance, yet it may not be in Usability. During a payment transaction, however, it must be ensured that the purchase is either completed or not-completed - halfway completed is not acceptable. for this it will be benificial to use the \emph{Rollback} tactic on the \emph{Basket} component.

@@

As it is not only the web server which may become non-responsive, it is also the database, it is also required to monitor the database. However, as a non-responsive data-base may be discovered by a non-responsive server, it is possible to use the ping-echo tactic with the database, and only perform the ping-echo if the server becomes non-responsive, to find out how much must be restarted (only the server or both the server and the database). 
Depending on the binding between teh database and the server, it may also be possibel to use the exceptions tactic between teh server and the database, so the server itself can restart the database, if an exception occur in a call to the database.

@@

As we have all experience the irritation of a slow web-site, and as the competition on the internet is growing, it is important to achieve propper performance.

Due to the lowing price of hardware and the increasing price of developement time, it is often cheapest and certainly simper to use the tactic \emph{Increase available resourcez}; introduce load balancing with multiple parallel servers thereby increase performance during heavy-load situations, ensure that the serves have sufficient network bandwidth, RAM and CPU strength to handle many simultanious database look-ups - here the tactic of \emph{Maintain multiple copies of data or computations} may be used to speed up database-look-up by caching the most popular searches, or cach the entire result in preperation for a search refinement. If the database is remotely located the caching of results becomes even more important.

There are many other tactics which may be employed, yet given the nature to the application, it is not believed that the biggest resource consumers are under developement control, for building the web-server from scratch to support better concurrency or using non-standard performance tweaking algorithms for the database-lookup, will most likely become inconcivably expensive compare to simply increasing the computational resources.

\fixme{Answer question 3.3}

\subsection{Quality Attributes - Question 4}

\begin{question}
Discuss where Service-Oriented Architecture can play a role
with respect to the Beer Web Store? Consider the quality implications
\end{question}

\fixme{Answer question 3.4}

>>>>

Consider the following main use case for a Beer Web Store for speciality
beers:

@@

Naturally in order for the scenarios to be able to be used to clarify the architecture it must be possible to quantify them, and place them in different categories. This requires that the above scenario(s) are extended with e.g. response time, how many times the user should be able to perform an operation before experiencing a fault (down-time).

Once the scenarios are clarified and voted on, then the top priority scenarios may be formalized and it may be determined which quality attribute(s) they involve. For examples sake lets assume that the above scenario was selected as top priority.

As it may be seen, the scenario(s) focus primarily on \emph{Availability} (how rare should it be that a user experience non-responsiveness and how many simultaneous users can the system support), \emph{Usability} (how easy should it be for the user to cancel a transaction halfway through and how easy should it be to pick up a previous transaction), \emph{Performance} (how big a delay can the user expect when performing an operation like a lookup in the database), \emph{Security} (how to avoid someone using stolen credit cards and also prevent a "hacker" from stealing the credit card information either during the transaction or from the Beer Web Store's database.

@@

The following scenarios (table \ref{tab:availability} and \ref{tab:performance}) is a possible outcome of prioritization and clarification of quality attribute scenarios as described above:

\begin{table}[H]
\begin{center}
\begin{tabular}{|p{0.3cm}|p{2.5cm}|p{8cm}|}
  \hline
  \multicolumn{2}{|p{3cm}|}{\bfseries Scenario(s):} & The customer browses the beer catalogue for descriptions and reviews and receives a timely response \\
  \hline
  \multicolumn{2}{|p{3cm}|}{\bfseries Relevant Quality Attributes:} & Availability, (Performance) \\
  \hline
  \multirow{6}{*}{\begin{sideways}{\bfseries Scenario Parts}\end{sideways}}
  & {\bfseries Source:} & The user \\
  \cline{2-3}
  & {\bfseries Stimulus:} & Browsing the beer catalogue \\
  \cline{2-3}
  & {\bfseries Artifact} & The system \\
  \cline{2-3}
  & {\bfseries Environment:} & During normal operation \\
  \cline{2-3}
  & {\bfseries Response:} & Web page presented to user \\
  \cline{2-3}
  & {\bfseries Response Measure:} & The result is presented within 5 seconds 99\% of the time, 30 seconds 0,9\% of the time and more than 30 seconds 0,1\% of the time \\
  \hline
  \multicolumn{2}{|p{3cm}|}{\bfseries Questions:} & \\
  \hline
  \multicolumn{2}{|p{3cm}|}{\bfseries Issues:} & The response measure is also dependent on the speed of the users internet connection and computer hardware. \\
  \hline
\end{tabular}
\caption{Availability scenario for the Beer Web Store}
\label{tab:availability}
\end{center}
\end{table}

\begin{table}[H]
\begin{center}
\begin{tabular}{|p{0.3cm}|p{2.5cm}|p{8cm}|}
  \hline
  \multicolumn{2}{|p{3cm}|}{\bfseries Scenario(s):} & When the customer checks out, payment is validated through the payment service within 30 seconds. \\
  \hline
  \multicolumn{2}{|p{3cm}|}{\bfseries Relevant Quality Attributes:} & Performance, (Availability) \\
  \hline
  \multirow{6}{*}{\begin{sideways}{\bfseries Scenario Parts}\end{sideways}}
  & {\bfseries Source:} & The user \\
  \cline{2-3}
  & {\bfseries Stimulus:} & The customer checks out \\
  \cline{2-3}
  & {\bfseries Artifact} & The system \\
  \cline{2-3}
  & {\bfseries Environment:} & During normal operation \\
  \cline{2-3}
  & {\bfseries Response:} & Result of payment validation if presented to the user \\
  \cline{2-3}
  & {\bfseries Response Measure:} & Payment is validated through the payment service within 30 seconds \\
  \hline
  \multicolumn{2}{|p{3cm}|}{\bfseries Questions:} & \\
  \hline
  \multicolumn{2}{|p{3cm}|}{\bfseries Issues:} & The response measure is also dependent on the performance of the payment service. \\
  \hline
\end{tabular}
\caption{Performance scenario for the Beer Web Store}
\label{tab:performance}
\end{center}
\end{table}

In a real life system there will of course be many more scenarios to describe the architectural requirements.

\subsection{Quality Attributes - Question 3}
\label{sec:qa_q3}

@@

Finding out if a resource is down may be done using one of the fault detection tactics; ping/echo, heartbeat or exception. Due to the loose coupling between the client (browser) and the server, the exception solution is not beneficial. Since it may be assumed that the supervising process (remotely deployed) (the process monitoring if the server is down) is stable, then a heartbeat is sufficient to ensure availability, meaning that the web-server will issue a heart-beat every e.g. 1 second to a maintenance process on the same or another machine, and if one (or perhaps more) heartbeats is missed then the maintenance service can act.

The proper response to a faulty server may be many things, as defined by the tactics group \emph{fault recovery}. As the requirement for availability stipulated a three stage fault recovery, the server or servers (as it may be required to have multiple serves to do load-balancing, depending on how many users must be supported) can service the 5 second requirement, and then if an error occur (as detected by the heartbeat) then restart or swap to another server/database in 30 seconds. As 30 seconds is insufficient to boot a web server and a database, the Spare tactic would be insufficient. At the same time having Active redundancy would be over-kill, as a this kind of immediate switch-over is not required. For these reasons Passive Redundancy is sufficient.

As it is an internet system, it will also be sufficient to simply restart the session, and no \emph{Shadow}, \emph{State resynchronization} or \emph{Rollback} is needed in the main web server. The user may experience being thrown back to the main page, yet this is deemed acceptable, under availability and performance, yet it may not be in Usability. During a payment transaction, however, it must be ensured that the purchase is either completed or not-completed - halfway completed is not acceptable. for this it will be beneficial to use the \emph{Rollback} tactic on the \emph{Basket} component.

@@

As it is not only the web server which may become non-responsive, it is also the database, it is also required to monitor the database. However, as a non-responsive data-base may be discovered by a non-responsive server, it is possible to use the ping-echo
 tactic with the database, and only perform the ping-echo if the server becomes non-responsive, to find out how much must be restarted (only the server or both the server and the database). 
Depending on the binding between the database and the server, it may also be possible to use the exceptions tactic between the server and the database, so the server itself can restart the database, if an exception occur in a call to the database.

@@

As we have all experience the irritation of a slow web-site, and as the competition on the internet is growing, it is important to achieve popper performance.

Due to the lowing price of hardware and the increasing price of development time, it is often cheapest and certainly simper to use the tactic \emph{Increase available resource}; introduce load balancing with multiple parallel servers thereby increase performance during heavy-load situations, ensure that the serves have sufficient network bandwidth, RAM and CPU strength to handle many simultanious database look-ups - here the tactic of \emph{Maintain multiple copies of data or computations} may be used to speed up database-look-up by caching the most popular searches, or cache the entire result in preparation for a search refinement. If the database is remotely located the caching of results becomes even more important.

There are many other tactics which may be employed, yet given the nature to the application, it is not believed that the biggest resource consumers are under development control, for building the web-server from scratch to support better concurrency or using non-standard performance tweaking algorithms for the database-lookup, will most likely become inconceivably expensive compare to simply increasing the computational resources.

\subsection{Quality Attributes - Question 4}
\label{sec:qa_q4}
\begin{question}
Discuss where Service-Oriented Architecture can play a role
with respect to the Beer Web Store? Consider the quality implications
\end{question}

As SOA is all about standardised interfaces supplying a functionality which may be used by many different products, the areas where SOA makes sense is naturally between the components.

Instead of creating a web-server which access a database through a specific fa\c ade, one may argue whether the desired functionality may not be generalised. "Look-up in database". This may easily be hidden behind a standardised service, e.g. a web-service, meaning that anyone on the WWW who supplies a database service may be used to host the beer database, and if one becomes dissatisfied with the database provider, it is transparent to simply switch to another.

Taking this one step further, perhaps the database service may be extended with a data-base type, meaning that the Beer Web Store's domain model is exposed in a service, which would allow for not only displaying ones own brands of beer, but through a discovery service, also beer-brands by other brewers on the net who implement this service. This will also point the way for a more standardised look-and-feel when shopping for beer.

Naturally just because a web-service (or other service) is used it does not mean that it has to be public, jet it does have some advantages.

Finally the WS-Security web-service may be used to handle payment, and also possibly, if the database or Beer Web Store Domain model is exposed through a web-service, it may also be beneficial that not everyone may have the ability to update the domain model or database, so here the WS-Security service may also come in handy.

Naturally this generality do not necessarily come free. A look-up in a database behind a service, especially a web-service, is far slower than a direct look-up in a local database, thereby influencing Performance poorly. On the other hand the use of web-services may increase Availability, as the functionality is deployed on multiple servers, thereby reducing the chance of the system becoming fully non-responsive. Naturally the critical path must be evaluated, as if a link on the critical path chain become non-responsive then the entire system becomes non-responsive.
